---
title: "SRT411 Assignment 0"
author: AJ
date: "January 29, 2017"
---

GITHUB USER: SRT411

The goal of Assignment 0 for SRT-411 Winter 2017 is to guide the student through the R programming language and introducing the fundamental and basic syntax of R-Studio and its various tools it comes with. Through this practical assignment we are guided through the process of creating a GitHub repository and using knitr creating a user friendly RMarkdown printout of our code.

Sources Used:

[Torfs Brauer Short R Intro](https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf)

[K Broman R Markdown](http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html)

[R-Studio R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

[K Broman Markdown](http://kbroman.org/knitr_knutshell/pages/markdown.html)

[Nicer Code](http://nicercode.github.io/guides/reports/)

[R Markdown R-Studio](http://rmarkdown.rstudio.com)

[Programiz If Else](https://www.programiz.com/r-programming/if-else-statement)

[Logical Operators](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Logic.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
###_**ToDo 1**_
Compute the difference between 2017 and the year you started at this university and divide this by the difference between 2017 and the year you were born. Multiply this with 100 to get the percentage of your life you have spent at this university. Use brackets if you need them.
```{r}
((2017-2015)/(2017-1990)*100)
```

***
###_**ToDo 2**_
Repeat the previous ToDo, but with several steps in between. You can give the variables any name you want, but the name has to start with a letter.
```{r}
startyear=2015
currentyear=2017
birthyear=1990
years=((currentyear-startyear)/(currentyear-birthyear))
totaltime=(years*100)
totaltime
```

***
###_**ToDo 3**_
Compute the sum of 4, 5, 8 and 11 by first com-bining them into a vector and then using the function sum.
```{r}
todo3=c(4,5,8,11)
sum(todo3)
```

***
###_**ToDo 4**_
Plot 100 normal random numbers.
```{r}
todo4=rnorm(100)
plot(todo4)
```

***
###_**ToDo 5**_
Find help for the sqrt function
```{r}
help(sqrt)
```

***
###_**ToDo 6**_
Make a file called firstscript.R containing R-code that generates 100 random numbers and plots them, and run this script several times.
```{r}
source("firstscript.R")
source("firstscript.R")
source("firstscript.R")
source("firstscript.R")
source("firstscript.R")
source("firstscript.R")
```

Contents of firstscript.R
```{r}
todo6 = rnorm(100)
plot(todo6)
```

***
###_**ToDo 7**_
Put the numbers 31 - 60 in a vector named P and in a matric with 6 rows and 5 columns named Q. Tip: use the function seq. Look at the different types of ways scalars, vectors and matrices are denoted in the workspace window.
```{r}
P=seq(from=31,to=60,by=1)
P
Q=matrix(data=P,nrow=6,ncol=5)
Q
```

***
###_**ToDo 8**_
Make a script file which constructs three ran- dom normal vectors of length 100. Call these vectors x1, x2 and x3. Make a data frame called t with three columns (called a, b and c) con- taining respectively x1, x1+x2 and x1+x2+x3. Call the following functions for this data frame: plot(t) and sd(t). Can you understand the results? Rerun this script a few times.
```{r}
source("secondscript.R")
source("secondscript.R")
source("secondscript.R")
source("secondscript.R")
source("secondscript.R")
source("secondscript.R")
```

Contents of secondscript.R
```{R}
x1=rnorm(100)
x2=rnorm(100)
x3=rnorm(100)

t=data.frame(a=x1,b=x1+x2,c=x1+x2+x3)

plot(t)
#sd() function no longer used, replaced with sapply()
sapply(t,sd)
```

***
###_**ToDo 9**_
Add these lines to the script file of the previous section. Try to find out, either by experiment- ing or by using the help, what the meaning is of rgb, the last argument of rgb, lwd, pch, cex.
```{r}
source("thirdscript.R")
```

Contents of thirdscript.R
```{R}
x1=rnorm(100)
x2=rnorm(100)
x3=rnorm(100)

t=data.frame(a=x1,b=x1+x2,c=x1+x2+x3)

plot(t)
#sd() function no longer used, replaced with sapply()
sapply(t,sd)

plot(t$a,type="l",ylim=range(t),lwd=3,col=rgb(1,0,0,0.3))
lines(t$b,type="s",lwd=2,col=rgb(0.3,0.4,0.3,0.9))
points(t$c,pch=20,cex=4,col=rgb(0,0,1,0.3))
```

_Meaning Of RGB/LWD/PCH/CEX:_

* rgb: red,green,blue colour specification

* lwd: line width for drawing symbols

* pch: the plotting 'character' to be used

* cex: the character/symbol expansion (numerical vector, default 1)

***
###_**ToDo 10**_

Make a file called tst1.txt in Notepad from the example in Figure 4 and store it in your working directory. Write a script to read it, to multiply the column called g by 5 and to store it as tst2.txt.
```{r}
source("fourthscript.R")
```

Contents of fourthscript.R
```{r}
readtodo10=read.table(file="tst1.txt",header=TRUE)
todo10=(readtodo10$g*5)
write.table(todo10,file="tst2.txt",row.names=FALSE)
t2=read.table(file="tst2.txt")
t2
```

Contents of tst1.txt

| a  | g  | x  |
|----|----|----|
| 1  | 2  | 3  |
| 2  | 4  | 6  |
| 4  | 8  | 12 |
| 8  | 16 | 24 |
| 16 | 32 | 48 |
| 32 | 64 | 96 |

Contents of tst2.txt

| x   |
|-----|
| 10  |
| 20  |
| 40  |
| 80  |
| 160 |
| 320 |

***
###_**ToDo 11**_
Compute the mean of the square root of a vec-tor of 100 random numbers. What happens?
```{r}
mean(sqrt(c(rnorm(100))))
```
_The operation returns NaNs, which means "Not a Number", this is because the square root of a negative integer does not exist, it will always return a NaN error when running this sort of operation, the mean function cannot return the average/mean of a non-integer value._

***
###_**ToDo 12**_
Make a graph with on the x-axis: today, Sinterklaas 2017 and your next birthday and on the y-axis the number of presents you expect on each of these days. Tip: make two vectors first.
```{r}
days=strptime(c("20170129","20171118","20180109"),format="%Y%m%d")
presents=c(0,0,4)
plot(days,presents,xlab="Days",ylab="Presents Expected")
```

***
###_**ToDo 13**_
Make a vector from 1 to 100. Make a for-loop which runs through the whole vector. Multiply the elements which are smaller than 5 and larger than 90 with 10 and the other elements with 0.1.
```{r}
todo13=seq(from=1,to=100,by=1)
for(i in 1:100)
{
	if (todo13[i] < 5 | todo13[i] > 90)
		{
		todo13[i]=(todo13[i]*10)	
		}
	else
	{
		todo13[i]=(todo13[i]*0.1)
	}
}
```

***
###_**ToDo 14**_
Write a function for the previous ToDo, so that you can feed it any vector you like (as argument). Use a for-loop in the function to do the computation with each element. Use the standard R function length in the specification of the counter.
```{r}
todo14=function(vector){
	for(i in 1:length(vector))
	{
		if (vector[i] < 5 | vector[i] > 90)
			{
			vector[i]=(vector[i]*10)	
			}
		else
			{
			vector[i]=(vector[i]*0.1)
			}
	}
	return(vector)
	}

example=seq(1:100)
todo14(vector=example)
```

***
###_**ToDo 15**_
Actually, people often use more for-loops than necessary. The ToDo above can be done more easily and quickly without a for-loop but with regular vector computations.
```{r}
todo15=seq(1:100)
todo15=ifelse(todo15 < 5 | todo15 > 90,todo15*10,todo15*0.1)
todo15
```
